---
title: "No 2. Data Preparation"
description: |
  Complete, reproducible workflow for preparation  of the **16S rRNA** data set. These steps are needed before analyzing the data. For aesthetic reasons, the R code for tables & figures are hidden (see the GitHub link at the bottom of the page for the raw `.Rmd` file). R code for everything else is generally shown.
bibliography: assets/cite.bib  
---

<details markdown="1">
<summary>Show setup information.</summary>

```{r setup, message = FALSE, warning = FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
set.seed(0199)
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
pacman::p_load(DT, ggplot2, dplyr, microbiome, tidyverse,   
               install = FALSE, update = FALSE)
options(scipen=999)
knitr::opts_current$get(c(
  "cache",
  "cache.path",
  "cache.rebuild",
  "dependson",
  "autodep"
))
```
</details>

```{r, echo=FALSE}
### NOTE: this should only be run AFTER the workflow is finished.
### This image is created at the END of this workflow and now read back in.
### Basically, this locks everything in place so nothing changes. Only code
### chunkc to display figures and tables are evaluated. Everything else is
### turned off. To actually run this you need to set this chunk to
### eval=FALSE and the rest to eval=TRUE
load("rdata/16s-data-prep/16s-data_prep.rdata")
```

# Data Availability

> You will either need to run the [DADA2 workflow](16s-dada2.html) or the grab output file `water_pipeline.rdata` from the workflow. This file contains the sequence and taxonomy tables. See the [Data Availability](data-availability.html) page for complete details.

All files needed to run this workflow can be downloaded from figshare. These are the output files from the [DADA2 workflow](16s-dada2.html) page.

<iframe src="https://widgets.figshare.com/articles/12403865/embed?show_title=1" width="568" height="300"  frameborder="0"></iframe>

File names and descriptions:

- **RUN01_read_changes.txt**: Tracking changes in read counts (per sample) from the beginning to end of the DADA2 workflow.
- **RUN02_read_changes.txt**: Tracking changes in read counts (per sample) from the beginning to end of the DADA2 workflow.
- **water_pipeline.rdata**: contains water sample sequence and taxonomy tables from the DADA2 pipeline needed for subsequent analyses. To see the Objects, in R run load("water_pipeline.rdata", verbose=TRUE)
- **combo_pipeline.rdata**: contains sequence and taxonomy tables from the DADA2 pipeline for all samples. To see the Objects, in R run load("combo_pipeline.rdata", verbose=TRUE)

`water_pipeline.rdata` contents. 

- **seqtab.1**: Sequence table from Run01 before merging with Run02.
- **seqtab.1**: Sequence table from Run02 before merging with Run01.
- **st.sum**: merged sequence table before removing chimeras
- **st.all**: duplicate of st.sum
- **seqtab**: merged sequence table after removing chimeras
- **tax_silva**: Silva (v132) taxonomy table of seqtab
- **tax_gg**: GreenGenes taxonomy table of seqtab

# Workflow Overview

Unless otherwise noted, we primarily use [phyloseq](https://joey711.github.io/phyloseq/) [@mcmurdie2013phyloseq] in this section of the workflow to analyze the 16S rRNA data set. Before we conduct any analyses we first need to *prepare our data set* by curating samples, removing *contaminants*, and creating phyloseq objects.

# Read Counts Assessment

Before we begin, let's create a summary table containing some basic sample metadata and the read count data from the [DADA2 pipeline](16s-dada2.html). We need to inspect how total reads changed through the pipeline. Just to get an idea, we combined the results of the [Track Changes](16s-dada2.html#track-read-changes) analysis for these 8 samples. Table headers are as follows:

| Header                    | Description                                                                              |
|---------------------------|------------------------------------------------------------------------------------------|
| `Type`                    | the type of source material for the sample (aka the habitat)                             |
| `Site`                    | the site where the sample was collected                                                  |
| `input`                   | number of raw reads after [`cutadapt`](16s-dada2.html#remove-primers)                    |
| `filter`                  | reads remaining after [QC & filtering](16s-dada2.html#quality-assessment-filtering)      |
| `denoiseF` & `denoiseR`   | reads after [error correction](16s-dada2.html#learn-error-rates)                         |
| `merged`                  | reads after [merging forward and reverse read](16s-dada2.html#merge-paired-reads)        |
| `nochim`                  | final read count after [removing chimeras](16s-dada2.html#remove-chimeras)               |
| `Change`                  | percent change from `input` to `nonchim`                                                 |

<br/>

```{r sample_table1, echo=FALSE, eval=FALSE}
joined_tab <- read.table(
  "tables/16s-data-prep/read_changes.txt",
  header = TRUE, sep = "\t")

joined_tab_percent <- joined_tab
percent_change <- 1-(joined_tab$nonchim/joined_tab$input) %>%
  round(digits = 3)
joined_tab_percent$Change <- as.numeric(sprintf("%.3f", percent_change))
colnames(joined_tab_percent) <- c("Sample<br/>ID", "Type", "Site",
                                  "input", "filter", "denoiseF", "denoiseR",
                                  "merged", "nochim", "Change")
```

```{r samp-summary1, echo=FALSE, layout="l-page"}
## NOTE. For some reason this datatable was being given the same
## elementId as the first table. So I had to add a 20 character
## elementId https://www.random.org/strings/
datatable(joined_tab_percent, width = "100%", escape = FALSE,
          rownames = FALSE, filter = 'top',
          caption = htmltools::tags$caption(
            style = 'caption-side: bottom; text-align: left;',
            'Table: ', htmltools::em('Tracking read changes
            through DADA2 workflow. Use the buttons to navigate through
            the table or download a copy. Table may scroll right. ')),
          elementId = "vwltiu9qj77sfy3a07we",
          extensions = 'Buttons', options = list(
            scrollX = TRUE,
            dom = 'Brti',
            buttons = c('copy', 'csv', 'excel'),
            pageLength = 8
            )
          ) %>%
  DT::formatStyle(columns = colnames(joined_tab_percent), fontSize = '80%')
cutoff <- 2000
remove_sam <- joined_tab[joined_tab$nonchim <= cutoff, ]
```

# Defining Groups

1. The first thing we want to do is load the data packet produced by the final step of the DADA2 workflow. This packet (`water_pipeline.rdata`) contains the ASV-by-sample table and the ASV taxonomy table.

2. After we load the data packet we next need to format sample names and define groups. We will use the actual sample names to define the different groups.

```{r deliniate_sample_types, eval=FALSE}
load("rdata/16s-dada2/water_pipeline.rdata")
samples.out <- rownames(seqtab)
subject <- sapply(strsplit(samples.out, "[[:digit:]]"), `[`, 1)
# this splits the string at first instance of a digit
sample_name <- substr(samples.out, 1, 999)  # use the whole string for individuals
type <- substr(samples.out, 0, 1)  # use the first two letters for sample typle
site <- substr(samples.out, 2, 3)  # use the next three letters for site
num_samp <- length(unique(sample_name))
num_type <- length(unique(type))
num_sites <- length(unique(site))
```

So  we have a total of **`r length(unique(sample_name))`** samples from **`r length(unique(site))`** sites.

> Sample abbreviations

The first letter of the sample ID indicates the environment:

W = Water

The next two letter indicates the site name:

CC = Coral Caye, CR = Cayo Roldan

An the number indicates the replicate number.

For example, `WCR3` is a `water` sample from  `Cayo Roldan` replicate `3`.

> Sample abbreviations:

* WCC = *Water sample from Coral Caye*
* WCR = *Water sample from Cayo Roldan*

4. And finally we define a sample data frame that holds the different groups we extracted from the sample names.

```{r define_variables}
#define a sample data frame
samdf <- data.frame(SamName = sample_name, TYPE = type, SITE = site)
rownames(samdf) <- samples.out
```

# Phyloseq Objects

**A**. The first step is to rename the amplicon sequence variants (ASVs) so the designations are a bit more user friendly. By default, DADA2 names each ASV by its unique sequence so that data can be directly compared across studies (which is great). But this convention can get cumbersome downstream, so we rename the ASVs using a simpler convention---ASV1, ASV2, ASV3, and so on.

<aside>
A phyloseq object contains ASV table (taxa abundances), sample metadata, and taxonomy table (mapping between ASVs and higher-level taxonomic classifications).
</aside>

```{r create_ps_object, eval=FALSE}
# this create the phyloseq object
ps <- phyloseq(otu_table(seqtab, taxa_are_rows = FALSE),
                   sample_data(samdf), tax_table(tax_silva))
tax_table(ps) <- cbind(tax_table(ps),
                           rownames(tax_table(ps)))

# adding unique ASV names
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
tax_table(ps) <- cbind(tax_table(ps),
                           rownames(tax_table(ps)))
head(taxa_names(ps))
```

```{r, echo=FALSE}
head(taxa_names(ps))
```

So the complete data set contains `r ntaxa(ps)` ASVs. We can also use the [microbiome R package](https://github.com/microbiome/microbiome/) [@lahti2017microbiome] to get some additional summary data from the phyloseq object.

```{r calculate_stats_ps, echo=FALSE, eval=FALSE}
min_read_ps <- min(readcount(ps))
max_read_ps <- max(readcount(ps))
total_reads_ps <- sum(readcount(ps))
mean_reads_ps <- round(mean(readcount(ps)), digits = 0)
median_reads_ps <- median(readcount(ps))
total_asvs_ps <- ntaxa(ps)
singleton_ps <- tryCatch(ntaxa(rare(ps, detection = 1, prevalence = 0)),
                              error=function(err) NA)
singleton_ps_perc <- tryCatch(round((100*(ntaxa(rare(ps, detection = 1, prevalence = 0)) /
                                   ntaxa(ps))), digits = 3), error=function(err) NA)
sparsity_ps <- round(length(which(abundances(ps) == 0))/length(abundances(ps)),
                     digits = 3)
```

| Metric                              | Results                                            |
|-------------------------------------|----------------------------------------------------|
| Min. number of reads                | `r min_read_ps`                                    |
| Max. number of reads                | `r max_read_ps`                                    |
| Total number of reads               | `r total_reads_ps`                                 |
| Average number of reads             | `r mean_reads_ps`                                  |
| Median number of reads              | `r median_reads_ps`                                |
| Sparsity                            | `r sparsity_ps`                                    |
| Any ASVs sum to 1 or less?          | `r isTRUE(singleton_ps >= 1)`                      |
| Number of singleton ASVs            | `r singleton_ps`                                   |
| Percent of ASVs that are singletons | `r singleton_ps_perc`                              |
| Number of sample variables are:     | `r length(sample_data(ps))`  (`r colnames(samdf)`) |

**B**. Add two final columns containing the ASV sequences and ASV IDs. This will be useful later when trying to export a fasta file. We can also take a look at the phyloseq object.

```{r add_ASV_coulmn, eval=FALSE}
colnames(tax_table(ps)) <- c("Kingdom", "Phylum", "Class", "Order",
    "Family", "Genus", "ASV_SEQ", "ASV_ID")
ps
```

```{r, echo=FALSE}
ps
```

**C**. Export sequence and taxonomy tables for the unadulterated phyloseq object for later use. We will use the prefix `full` to indicate that these are the *raw* outputs.

```{r export_seq_tax_tables, eval=FALSE}
write.table(tax_table(ps),
            "tables/16s-data-prep/full_tax_table.txt",
            sep="\t", quote = FALSE, col.names=NA)
write.table(t(otu_table(ps)),
            "tables/16s-data-prep/full_seq_table.txt",
            sep="\t", quote = FALSE, col.names=NA)
```

**D**. Export fasta file

```{r export_fasta_full, eval=FALSE}
# Create fasta file from tax_table
table2format <- tax_table(ps)
#retain only the column with the sequences
table2format_trim <- table2format[, 7]
table2format_trim_df <- data.frame(row.names(table2format_trim),
                                   table2format_trim)
colnames(table2format_trim_df) <- c("ASV_ID", "ASV_SEQ")
#format fasta
table2format_trim_df$ASV_ID <- sub("ASV", ">ASV", table2format_trim_df$ASV_ID)

write.table(table2format_trim_df, "tables/16s-data-prep/full_asv.fasta",
            sep = "\r", col.names = FALSE, row.names = FALSE,
            quote = FALSE, fileEncoding = "UTF-8")
```

# Curation

Let's see if we have any potential contaminants. We can use some [inline R code](https://rmarkdown.rstudio.com/lesson-4.html) to see the taxonomy table for any taxa of interest.

<aside>
The code (hidden by default) is written as `` `r
"Chloroplast" %in% tax_table(ps)` ``.
</aside>

- Are Mitochondria present? `r "Mitochondria" %in% tax_table(ps)`
- Are Chloroplast present? `r "Chloroplast" %in% tax_table(ps)`
- Are Eukaryota present? `r "Eukaryota" %in% tax_table(ps)`

Let's remove these taxa---Eukaryota because we used bacterial/archaeal primers, Mitochondria because those are likely from  eukaryotes, and Chloroplast because those are likely from algae. We must do each of these in turn using phyloseq and it gets a little messy.

Why messy? The `subset_taxa` command removes anything that is `NA` for the specified taxonomic level or above. For example, let’s say you run the `subset_taxa` command using `Family != "Mitochondria`". Seems like you should get a phyloseq object with everything except Mitochondria. But actually, the command not only gets rid of Mitochondria but everything else that has `NA` for Family and above. In my experience this is not well documented, and I had to dig through the files to figure out what was happening.

Anyway, to remove the taxa we do the following:

* Subset the taxa and generate a `ps` object of just the taxa of interest,
* Select the ASV column only, turn it into a factor, and use this to remove <INSERT TAXA> from the `ps` object.

## Remove  Mitochondria ASVs

Remember the original data set contained `r ntaxa(ps)` ASVs.

```{r remove_specific_taxa, eval=FALSE}
# generate a file with mitochondria ASVs
MT1 <- subset_taxa(ps, Family == "Mitochondria")
MT1 <-  as(tax_table(MT1), "matrix")
MT1 <- MT1[, 8]
MT1df <- as.factor(MT1)
goodTaxa <- setdiff(taxa_names(ps), MT1df)
ps_no_mito <- prune_taxa(goodTaxa, ps)
ps_no_mito
```

```{r, echo=FALSE}
ps_no_mito
```

Looks like this removed **`r ntaxa(ps) - ntaxa(ps_no_mito)` Mitochondria ASVs**. We will duplicate the code block to remove other groups.

## Remove Chloroplast ASVs

```{r remove_specific_taxa2, eval=FALSE}
# generate a file with mitochondria ASVs
CH1 <- subset_taxa(ps_no_mito, Order == "Chloroplast")
CH1 <-  as(tax_table(CH1), "matrix")
CH1 <- CH1[, 8]
CH1df <- as.factor(CH1)
goodTaxa <- setdiff(taxa_names(ps_no_mito), CH1df)
ps_no_chloro <- prune_taxa(goodTaxa, ps_no_mito)
ps_no_chloro
```

```{r, echo=FALSE}
ps_no_chloro
```

The code removed an additional **`r ntaxa(ps_no_mito) - ntaxa(ps_no_chloro)` Chloroplast ASVs**.

## Remove Eukaryotic ASVs

Turns out we had no eukaryotic ASVs but here is the code just in case.

```{r remove_specific_taxa3, eval=FALSE, eval=FALSE}
# generate a file with mitochondria ASVs
EU1 <- subset_taxa(ps_no_chloro, Kingdom == "Eukaryota")
EU1 <-  as(tax_table(EU1), "matrix")
EU1 <- EU1[, 8]
EU1df <- as.factor(EU1)
goodTaxa <- setdiff(taxa_names(ps_no_chloro), EU1df)
ps_no_euk <- prune_taxa(goodTaxa, ps_no_chloro)
ps_no_euk
```

## Remove any Kingdom NAs

Here we can just use the straight up `subset_taxa` command since we do not need to worry about any ranks above Kingdom also being removed.

```{r remove_kingdom_na, eval=FALSE}
ps_filt <- subset_taxa(ps_no_chloro, !is.na(Kingdom))
ps_filt
```

```{r, echo=FALSE}
ps_filt
```

The code eliminated an additional **`r ntaxa(ps_no_chloro) - ntaxa(ps_filt)` Kingdom level NA ASVs** from the phyloseq object.

Now that we have a trimmed data set, we can save copies of the sequence and taxonomy tables just in case we want a quick look at these data.

```{r save_trim_table, eval=FALSE}
write.table(tax_table(ps_filt),
            "tables/16s-data-prep/trim_tax_table.txt", sep="\t",
            quote = FALSE, col.names=NA)
write.table(t(otu_table(ps_filt)),
            "tables/16s-data-prep/trim_seq_table.txt", sep="\t",
            quote = FALSE, col.names=NA)
write.table(sample_data(ps_filt),
            "tables/16s-data-prep/sample_table.txt",
            sep="\t", quote = FALSE, col.names=NA)
```

And generate a fasta file for all ASVs in the trimmed data set.

```{r export_fasta_water, eval=FALSE}
# Create fasta file from tax_table
table2format <- tax_table(ps_filt)
#retain only the column with the sequences
table2format_trim <- table2format[, 7]
table2format_trim_df <- data.frame(row.names(table2format_trim),
                                   table2format_trim)
colnames(table2format_trim_df) <- c("ASV_ID", "ASV_SEQ")
#format fasta
table2format_trim_df$ASV_ID <- sub("ASV", ">ASV", table2format_trim_df$ASV_ID)

write.table(table2format_trim_df, "tables/16s-data-prep/trim_asv.fasta",
            sep = "\r", col.names = FALSE, row.names = FALSE,
            quote = FALSE, fileEncoding = "UTF-8")
```


```{r select_water_samples, echo=FALSE, eval=FALSE}
###############################################
## This is legacy code for making a ps object##
## for a subset of samples                   ##
###############################################

## Water Samples Phyloseq Object

#The next thing to do is select only water samples and make a separate phyloseq object---since our paper is about the water samples. To do this we must select the samples we *wish to keep*. If you want to change the group of samples, modify the script accordingly.

ps_water <- prune_samples(
  c("WCC0", "WCC1", "WCC2", "WCC3", "WCR0", "WCR1", "WCR2", "WCR3"),
  ps_filt)
ps_water

#Anytime we remove samples we probably lose some ASVs. So we need to get rid of any ASVs that have a total of **0 reads**.
ps_water <- prune_taxa(taxa_sums(ps_water) > 0, ps_water)
ps_water

#So for water samples only there are `r ntaxa(ps_water)` ASVs and `r nsamples(ps_water)` samples. Now, save the various phyloseq data from the water object.
```

# Save Objects

One last thing is to rename the phyloseq object and save some tables and images for the next part.

```{r save_data_prep_image, echo=TRUE, eval=FALSE}
ps_water <- ps_filt
save.image("rdata/16s-data-prep/16s-data_prep.rdata")
saveRDS(otu_table(ps_water), "rdata/16s-data-prep/ps_water-seqtab.rds")
saveRDS(sample_data(ps_water), "rdata/16s-data-prep/ps_water-sample.rds")
saveRDS(tax_table(ps_water), "rdata/16s-data-prep/ps_water-taxtab.rds")
```

That's all for this part! In the next section we work through analyzing the diversity of the water samples.

</br>

<div class="post-nav">
<div class="post-nav-item">
<div class="meta-nav">Previous</div>
<a href="16s-dada2.html" rel="next">N<sup><u>o</u></sup> 1. DADA2 Workflow</a>
</div>
</div>
<div class="post-nav">
<div class="post-nav-item">
<div class="meta-nav">Next</div>
<a href="16s-water.html" rel="prev">N<sup><u>o</u></sup> 3. Water Composition & Diversity</a>
</div>
</div>

## Source Code {.appendix}

You can find the source code for this page by [clicking this link](https://github.com/hypocolypse/web/blob/master/16s-data-prep.Rmd).


## Data Availability {.appendix}

Output files from this workflow available on figshare at [doi:10.25573/data.12403925](https://doi.org/10.25573/data.12403925). Here you will find the sample metadata table, sequence tables, taxonomy tables, and ASV fasta files from the full phyloseq object (before removing contaminants) and the trimmed data set (after removing contaminants). The *16s-data_prep.rdata* file contains all variables and phyloseq objects from this pipeline and is needed for the next workflow (as are the .rds files).
