---
title: "Field Analyses"
description: |
 In this section we discuss the various field analyses conducted in this study.
output:
    distill::distill_article:
      css: assets/styles.css
      toc: true
      toc_depth: 3
bibliography: assets/cite.bib  
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(lme4)
library(lmerTest)
library(car)
library(rcompanion)
library(MASS)
library(emmeans)
library(nlme)
library(ggplot2)
library(plyr)
library(RColorBrewer)
library(wesanderson)
library(reshape2) # restructuring the data
library(cowplot)
library(scales)
library(tidyr)
library(dplyr)
library(patchwork)
```


## Hypoxic Reef Conditions

Conditions characteristic of a hypoxic event were first detected on Sept. 20, 2017 in Bahia Almirante on the Caribbean coast of Panama. To evaluate the environmental conditions associated with the event we conducted a series of depth profiles with a YSI multiparameter sonde, measuring dissolved oxygen (DO), temperature, chlorophyll, salinity, and pH. Measurements were taken at 83 sites across Bahía Almirante on Sept. 25, 2017. We selected two focal sites across a gradient of DO concentrations, from extremely hypoxic to normoxic (ambient DO concentrations). Cayo Roldan (hypoxic site) lies at the inner reaches of Bahia Almirante where water exchange with the open ocean is restricted, experience extreme hypoxia on the benthos during the event. Cayo Coral (normoxic site) is well-flushed with open ocean water and was used as a control site because oxygenated conditions persisted during the event. 

## Coral Bleaching

The common lettuce coral, *Agaricia tenuifolia*, was collected from each site during the hypoxic event and evaluated for bleaching through measurements of symbiont densities, pigment content, and pulse amplitude modulated (PAM) fluorometry. Three fragments from each of six colonies were collected from each site and returned to the laboratory facilities at the Smithsonian Tropical Research Institute's Bocas del Toro Research Station. Differences in each response variable were evaluated with a linear mixed effects model, with colony as a random effect and size as a fixed effect. 

## Coral Community Structure

The hypoxic and normoxic sites are the location of permanent benthic monitoring. We evaluated changes in benthic community structure from four months before, during, and one year after the hypoxic event using permanent photoquadrats. At each site, permanent quadrats were photographed every 5m along one 50m permanent transect, and community structure was determined from 50 randomly overlaid points using the online platform [CoralNet](https://coralnet.ucsd.edu/). Taxa were identified to the finest resolution possible, and then grouped into different functional groups for subsequent analyses. 

We used nonmetric multidimensional scaling (nMDS) to visually represent benthic community structure at each time point, and evaluated changes over time with a permutational analysis of variance (PERMANOVA), with pairwise contrasts. The benthic community did not change significantly over time at the normoxic site, so changes in the coral community over time were evaluated only at the hypoxic site. 

## Analysis Workflow 

Next, we present the workflows we used to analyze coral response variables, specifically the Pulse Amplitude Modulated (PAM) Fluorometry, Zooxanthellae, Chlorophyll a (Chla), and Chlorophyll C (Chlc) data. For each data set, we visually assess residuals, test for normality (Shapiro-Wilk Test), and test for equality of variances (Levene’s Test). In some cases, data are subsequently transformed. 

```{r}
data <- read.table("tables/field/coralhypoxia.csv", header = T, sep = ",")
attach(data)
```

## Pulse Amplitude Modulated (PAM) Fluorometry

First, we test all data for assumptions of normality and homogeneity of variances based on raw data (i.e., not transformed) and residuals. Data are normal, variances hetergeneous (transform doesn't improve). Look at distribution of data with histogram.

### Non-Transformed Data

#### Visual Assessment

```{r}
plotNormalHistogram(yield, main = "PAM", xlab = "Yield")
```

First, we run the model to generate the residuals, then look at residuals to see if they are normally distributed. This is a mixed model with fragment nested in colony and site as the fixed effect.

```{r}
lmepam <- lmer(yield ~ site + (1|colony), REML = FALSE, na.action = na.exclude)
res1 <- resid(lmepam)
```

And then produce a histogram of the residuals.

```{r, eval=FALSE, include=FALSE}
hist(res1, main = "PAM", xlab = "Residuals")
plotNormalHistogram(res1)
```

```{r}
plotNormalHistogram(res1, main = "PAM", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

Next, we run a normality test on residuals. The null hypothesis is that the data is normally distributed, while a p-value < 0.05 means the data is non-normally distributed.

```{r}
pam.shapiro <- shapiro.test(res1)
print(pam.shapiro)  
```

Ok, the p-value is `r pam.shapiro$p.value`, meaning the result is not significant and the data are normally distributed.

Next, we generate QQ plots of residuals from the lme model.

```{r}
qqnorm(res1, ylab = "Sample Quantiles for PAM")
qqline(res1, col = "red")
```

#### Levene’s Test

And the Levene’s Test, which tests for homogeneity of variance across groups.

```{r}
leveneTest(yield ~ as.factor(site), data = data)
bartlett.test(x = yield, g = site)
```

So the data is ***normal but not homogenous***.

#### Mixed Effects Models

In our mixed model approach, **site** is the *fixed* effect, **colony** the *random* effect, and **fragment** is random nested in **colony**. We compared the full model with fragment dropped. The best fit model (with lowest AIC) was the model with colony dropped. But we use a model that includes fragment because it's a necessary random effect.

When comparing models with different random effects like here---use `REML = FALSE`. But when not comparing use `REML = TRUE` (default). The full model includes site, colony, and fragment nested in colony. Colony/frag includes a term for colony alone and nests fragment in colony.

```{r}
pam1 <- lmer(yield ~ site + (1|colony/frag), data = data, 
             REML = FALSE, na.action = na.exclude)
ranova(pam1)
```

So we drop fragment nested in colony.

```{r}
pam2 <- lmer(yield ~ site + (1|colony), data = data, 
             REML = FALSE, na.action = na.exclude)
ranova(pam2)
```
And then use the full model to account for nesting fragments.

```{r}
mod.AIC <- AIC(pam1, pam2) 
mod.AIC 
```

And the final models use REML but without comparing models with different random effects.

```{r}
finalpam <- lmer(yield ~ site + (1|colony), 
                 data = data, 
                 REML = TRUE, na.action = na.exclude)
```

p values for random effects.

```{r}
finalrandpam <- ranova(finalpam)
pam.table <- anova(finalpam, type = 2)
```

```{r}
pam.table
```

```{r}
finalrandpam
```

## Zooxanthellae  

First, we test all data for assumptions of normality and homogeneity of variances. Based on the raw data and residuals, CR6B is an outlier and was removed. Log-transforming the data makes the data normal and homogeneous. We will look at both non-transformed transformed data.

### Non-Transformed Data

Look at distribution of the raw data with histogram.

```{r}
attach(data)
plotNormalHistogram(zoox, main = "Zooxanthellae", xlab = "Counts")
```

First, run the model to generate the residuals, then look at residuals to see if thay are normally distributed. Mixed model with fragment nested in colony, site as fixed effect. Yield is normally distributed, but variances not homogeneous.

```{r}
lmezoox <- lmer(zoox ~ site + (1|colony), REML = FALSE, na.action = na.exclude)
res2 <- resid(lmezoox)
```

Produce a histogram of the residuals.

```{r, include=FALSE, eval = FALSE}
hist(res2, main = "Zoox", xlab = "Residuals")
plotNormalHistogram(res2, main = "Zooxanthellae", xlab = "Residuals")
```

```{r}
plotNormalHistogram(res2, main = "Zooxanthellae", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

Next, we again run a normality test on residuals. And again, the null hypothesis is that the data is normally distributed, while a p-value < 0.05 means the data is non-normally distributed.

```{r}
zoox.shapiro <- shapiro.test(res2) #runs a normality test on residuals
print(zoox.shapiro) # null = normally distrubuted (P<0.05 = non-normal
```
We see the p-value (`r zoox.shapiro$p.value`) is less than 0.05, and thus the data is non-normally distributed. 

And a quick look at the QQ plot of residuals from lme model.

```{r}
qqnorm(res2, ylab = "Sample Quantiles for Zoox")
qqline(res2, col = "red")
```

#### Levene’s Test

```{r}
leveneTest(zoox~ as.factor(site),data = data) # p = 0.0007486
```

### Transformed Data

Log transformation improves normality. First, we add a new column to data frame for the log transformation and check that log transformed data is added as a new column.

```{r}
data$zoox_log = log(zoox)
attach(data)
colnames(data)
```

#### Visual Assessment

Check normality of log-transformed data.

```{r}
plotNormalHistogram(zoox_log, main = "Zooxanthellae", xlab = "log (counts)")
```

Check normality of log-transformed residuals

```{r}
lmezoox_log <- lmer(zoox_log ~ site + (1|colony), 
                    REML = FALSE, 
                    na.action = na.exclude)
res3 <- resid(lmezoox_log)
```

Produce a histogram of the residuals.

```{r, include=FALSE, eval = FALSE}
hist(res3, main = "zoox log", xlab = "Residuals")
plotNormalHistogram(res3, main = "Zooxanthellae log", xlab = "Residuals")
```

```{r}
plotNormalHistogram(res3, main = "Zooxanthellae log", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

We then run a normality test on residuals. 

```{r}
zooxlog.shapiro <- shapiro.test(res3)
print(zooxlog.shapiro)
```

Now the p-value (`r zooxlog.shapiro$p.value`) is greater than 0.05, indicating the log-transformed data is normally distributed.

QQ plots of residuals from lme model.

```{r}
qqnorm(res3, ylab = "Sample Quantiles for Zooxlog")
qqline(res3, col = "red")
```

#### Levene's Test

```{r}
leveneTest(zoox_log ~ as.factor(site), data = data)
```

#### Mixed Effects Models

For the Mixed Effects Models, site is a fixed effect, colony is random, and fragment is random nested in colony. Colony/fragment includes a term for colony alone and nests fragment in colony.

```{r}
zoox1 <- lmer(zoox_log ~ site + (1|colony/frag), 
              data = data, 
              REML = FALSE, 
              na.action = na.exclude) #AIC 72.8756
ranova(zoox1)
```

Fragment nested in colony is then dropped.

```{r}
zoox2 <- lmer(zoox_log ~ site + (1|colony), 
              data = data, 
              REML = FALSE, 
              na.action = na.exclude)
ranova(zoox2)
mod.AIC <- AIC(zoox1, zoox2) 
mod.AIC
```

The Ffinal models uses REML but not comparing models with different random effects.

```{r}
finalzoox <- lmer(zoox_log ~ site + (1|colony), 
                  data = data, 
                  REML = TRUE, 
                  na.action = na.exclude)
```

P-values for random effects

```{r}
finalrandzoox <- ranova(finalzoox)
zoox.table <- anova(finalzoox, type = 2) #ANOVA table
zoox.table
finalrandzoox
```

## Chlorophyll a (Chla)

As before, we first test all data for assumptions of normality and homogeneity of variances. 

### Non-Transformed Data

#### Visual Assessment

Look at distribution of the raw data with histogram.

```{r}
plotNormalHistogram(chla, main = "Chlorophyll a", xlab = "Chlorophyll")
```

First run the model to generate the residuals, then look at residuals to see if normally distributed. Mixed model with fragment nested in colony, site as fixed effect. Raw data are normal but variances not homogenous---log-transformed improves homogeneity.

```{r}
lmechla <- lmer(chla ~ site + (1|colony), 
                REML = FALSE, 
                na.action = na.exclude)
res4 <- resid(lmechla)
```

Produce a histogram of the residuals.

```{r, include=FALSE, eval=FALSE}
hist(res4, main = "Chla", xlab = "Residuals")
plotNormalHistogram(res4, main = "Chlorophyll a", xlab = "Residuals")
```


```{r}
plotNormalHistogram(res4, main = "Chlorophyll a", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

```{r}
chla.shapiro <- shapiro.test(res4)
print(chla.shapiro) 
```

QQ plots of residuals from lme model

```{r}
qqnorm(res4, ylab = "Sample Quantiles for Zoox")
qqline(res4, col = "red")
```

#### Levene's Test

```{r}
leveneTest(chla ~ as.factor(site), data = data) # p = 0.01606
```

Data are normal but variances are not homogenous.

### Transformed Data

Add a new column to dataframe for log transformation.

```{r}
data$chla_log = log(chla)
```

Check that log transformed data is added as a new column

```{r}
head(data)
attach(data)
colnames(data)
```

#### Visual Assessment

Check normality of log-transformed data.

```{r}
plotNormalHistogram(chla_log, main = "Chlorophyll a", xlab = "log (Chlorophyll)")
```

Check normality of log-transformed residuals.

```{r}
lmechla_log <- lmer(chla_log ~ site + (1|colony), 
                    REML = FALSE, 
                    na.action = na.exclude)
res7 <- resid(lmechla_log)
```

Produce a histogram of the residuals.

```{r, include=FALSE, eval = FALSE}
hist(res7, main = "chla log", xlab = "Residuals")
plotNormalHistogram(res7, main = "Chlorophyll a", xlab = "Residuals")
```


```{r}
plotNormalHistogram(res7, main = "Chlorophyll a", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

```{r}
chlalog.shapiro <- shapiro.test(res7) #runs a normality test on residuals
print(chlalog.shapiro) # null = normally distrubuted (P<0.05 = non-normal
```

QQ plots of residuals from lme model.

```{r}
qqnorm(res7, ylab = "Sample Quantiles for chlalog")
qqline(res7, col = "red")
```

#### Levene's Test

```{r}
leveneTest(chla_log ~ as.factor(site), data = data) # p = 0.7487
```

#### Mixed Model Effects

Site is fixed effect, colony is random, fragment is random nested in colony on log transformed data. Colony/frag includes a term for colony alone and nests fragment in colony. 

```{r}
chla_log1 <- lmer(chla_log ~ site+(1|colony/frag), 
                  data = data,
                  REML = FALSE, 
                  na.action = na.exclude) #AIC 55.11567
ranova(chla_log1)
```

Fragment nested in colony dropped.

```{r}
chla_log2 <- lmer(chla_log ~ site + (1|colony), 
                  data = data, 
                  REML = FALSE, 
                  na.action = na.exclude)
ranova(chla_log2)

mod.AIC <- AIC(chla_log1, chla_log2) 
mod.AIC
```

Final models uses REML---not comparing models with different random effects.

```{r}
finalchla_log <- lmer(chla_log ~ site + (1|colony), 
                      data = data, 
                      REML = TRUE, 
                      na.action = na.exclude)
```

p values for random effects.

```{r}
finalchla_logrand <- ranova(finalchla_log)
chla_log.table <- anova(finalchla_log, type = 2) #ANOVA table
chla_log.table
finalchla_logrand
```

## Chlorophyll c (Chlc)

First testing all data for assumptions of normality and homogeneity of variances. Based on raw data and residuals. Data need to be log transformed for normality and variances.

### Non-Transformed Data

#### Visual Assessment

Look at distribution of the raw data with histogram.

```{r}
plotNormalHistogram(chlc, main = "Chlorophyll c", xlab = "Chlorophyll")
```

Run the model to generate the residuals then look at residuals to see if normally distributed mixed model with fragment nested in colony, site as fixed effect yield is normally distributed, but variances not homogeneous.

```{r}
lmechlc <- lmer(chlc ~ site + (1|colony), 
                REML = FALSE, 
                na.action = na.exclude)
res5 <- resid(lmechlc)
```

Produce a histogram of the residuals.

```{r, eval=FALSE, include=FALSE}
hist(res5,main = "Chlorophyll c", xlab = "Residuals")
plotNormalHistogram(res5, main = "Chlorophyll c", xlab = "Residuals")
```

```{r}
plotNormalHistogram(res5, main = "Chlorophyll c", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

```{r}
chlc.shapiro <- shapiro.test(res5)
print(chlc.shapiro)
```

QQ plots of residuals from lme model.

```{r}
qqnorm(res5, ylab = "Sample Quantiles for Zoox")
qqline(res5, col = "red")
```

#### Levene's Test

```{r}
leveneTest(chlc ~ as.factor(site), data = data) # p = 0.01316
```

Data not normally distributed or homogenous

### Transformed Data

```{r}
data$chlc_log <- log(chlc)
attach(data)
colnames(data)
```

#### Visual Assessment

Check normality of log-transformed data

```{r}
plotNormalHistogram(data$chlc_log, main = "Chlorophyll c (log)", xlab = "log (Chlorophyll)")
```

Check normality of log-transformed residuals.

```{r}
lmechlc_log <- lmer(chlc_log ~ site + (1|colony), 
                    REML = FALSE, 
                    na.action = na.exclude)
res6 <- resid(lmechlc_log)
```

Produce a histogram of the residuals.

```{r, include=FALSE, eval=FALSE}
hist(res6, main = "Chlorophyll c log", xlab = "Residuals")
plotNormalHistogram(res6, main = "Chlorophyll c log", xlab = "Residuals")
```

```{r}
plotNormalHistogram(res6, main = "Chlorophyll c (log)", xlab = "Residuals")
```

#### Shapiro-Wilk Normality Test 

```{r}
chlclog.shapiro <- shapiro.test(res6)
print(chlclog.shapiro)
```

QQ plots of residuals from lme model.

```{r}
qqnorm(res6, ylab = "Sample Quantiles for chlclog")
qqline(res6, col = "red")
```

#### Levene's Test

```{r}
leveneTest(chlc_log ~ as.factor(site), data = data) # p = 0.884
```

#### Mixed Model Effects

Site is the fixed effect, colony is random, fragment is random nested in colony. On log-transformed data. Colony/fragment includes a term for colony alone and nests fragment in colony.

```{r}
chlc_log1 <- lmer(chlc_log ~ site + (1|colony/frag), 
                  data = data, 
                  REML = FALSE, 
                  na.action = na.exclude) #AIC 64.52399
ranova(chlc_log1)
```

Fragment nested in colony dropped.

```{r}
chlc_log2 <- lmer(chlc_log ~ site+(1|colony), 
                  data = data, 
                  REML = FALSE, 
                  na.action = na.exclude)
ranova(chlc_log2)

mod.AIC <- AIC(chlc_log1, chlc_log2) 
mod.AIC
```

Final models use REML.

```{r}
finalchlc_log <- lmer(chlc_log ~ site + (1|colony), 
                      data = data, 
                      REML = TRUE, 
                      na.action = na.exclude)
```

P-values for random effects.

```{r}
finalchlc_logrand <- ranova(finalchlc_log)
chlc_log.table <- anova(finalchlc_log, type = 2) #ANOVA table
chlc_log.table
finalchlc_logrand
```

## Figures

```{r, }
all <- read.table("tables/field/coralhypoxia.csv", header = T, sep = ",")
attach(all)
```

### Calculate Means for Figures 

```{r}
means.all <- ddply(all, c("site", "colony"), summarise,
                 N    = sum(!is.na(yield)),
                 yield_mean = mean(yield, na.rm = TRUE),
                 zoox_mean = mean(zoox, na.rm = TRUE),
                 chla_mean = mean(chla, na.rm = TRUE),
                 chlc_mean = mean(chlc, na.rm = TRUE),
                 sd_yield = sd(yield),
                 sd_zoox = sd(zoox),
                 sd_chla = sd(chla),
                 sd_chlc = sd(chlc),
                 yield_se = (sd(yield)/sqrt(N)),
                 zoox_se = (sd(zoox)/sqrt(N)),
                 chla_se = (sd(chla)/sqrt(N)),
                 chlc_se = (sd(chlc)/sqrt(N)))
write.csv(means.all, file = "tables/field/allcoralmeans.csv", 
          row.names = FALSE)
```

### PAM

```{r}
ggplot(all,aes(x = site, y = yield)) +
  geom_boxplot() +
  labs(x = "Site", y='Maximum Quantum Yield') + 
  theme_classic() +
  theme(text = element_text(size=14)) +
  theme(axis.text.x = element_text(colour = "black", angle = 0, hjust = 1),
        axis.text.y = element_text(colour = "black")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0.4,0.75))
```

```{r}
ggplot(all,aes(x=site, y=yield, color=colony)) +
  geom_point(size=3) + 
  labs(x = "Site", y = 'Maximum Quantum Yield') + 
  theme_classic() +
  theme(text = element_text(size=14)) +
  theme(axis.text.x = element_text(colour="black", angle=0, hjust=1),
        axis.text.y = element_text(colour="black")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0.4,0.75))
```

```{r}
ggplot(means.all,aes(x = site, y = yield_mean,fill = site)) +
  geom_boxplot() +
  labs(x = "Site", y = 'Maximum Quantum Yield') + 
  theme_classic() +
  scale_fill_manual(values = c("steelblue","darkred")) +
  theme(text = element_text(size = 14)) + 
  theme(axis.text.x = element_text(colour = "black", angle = 0, hjust = 1),
        axis.text.y = element_text(colour = "black")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0.4,0.75)) 
```

Set a jitter object to offset points and assign data to ggplot.

```{r}
jitter <- position_jitter(width = 0.1)  
yield_plot <- ggplot(means.all, aes(x = site, y = yield_mean))
```

```{r}
yp <- yield_plot+
  geom_boxplot(color="black",
               aes(fill=site),
               width=0.4) +
  scale_fill_manual(values=c("skyblue3","coral3"), 
                    name= NULL, labels=c("Normoxic","Hypoxic")) +
  geom_point(size=2, position=jitter) +
  labs(x="Site",y="Maximum Quantum Yield") +
  theme_classic() +
  theme(legend.position = c(0.85,0.85),
        legend.title= element_text(color="transparent"),
        axis.text.x = element_text(colour="black"), 
        axis.text.y = element_text(colour="black"),
        axis.title.y = element_text(margin = margin(t=0, r=10,b=0,l=0)),
        panel.border = element_rect(colour = "black", fill=NA, size=.5)) +
  theme(text = element_text(size=12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay","Roldan" = "Cayo Roldan")) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,0.7))
yp
```

### Zooxanthellae Counts

Assign data to ggplot.

```{r}
yield_zoox <- ggplot(means.all, aes(x=site, y=zoox_mean))
zp <- yield_zoox+
  geom_boxplot(color="black",
               aes(fill=site),
               width=0.4) +
  scale_fill_manual(values=c("skyblue3","coral3"), 
                    name= NULL, labels=c("Normoxic","Hypoxic")) +
  geom_point(size=2, position=jitter) +
  labs(x="Site",
       y=("Endosymbiont Density (cells cm^-2)")) +
  theme_classic() +
  theme(legend.position = "none", #
        axis.text.x = element_text(colour="black"),
        axis.text.y = element_text(colour="black"),
        axis.title.y = element_text(margin = margin(t=0, r=10,b=0,l=0)),
        panel.border = element_rect(colour = "black", fill=NA, size=.5)) +
  theme(text = element_text(size=12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay","Roldan" = "Cayo Roldan")) +
  scale_y_continuous(labels = scientific,
                     breaks=seq(0,1.5E6, 4.5E5),
                      expand = c(0,0), 
                     limits = c(0,1.5E6))
zp  
```

### Pigment Data 

Convert Pigment data to long format with Chlorophyll a and Chlorophyll c stacked.

```{r}
pigs <- tidyr::gather(means.all, pigment, concentration,chla_mean, chlc_mean)
pigments <- dplyr::select(pigs,site,colony,pigment,concentration)
```

Assign data to ggplot data frame "pigments".

```{r}
pig_plot <- ggplot(pigments,
                     aes(x=pigment, y=concentration, fill=site))
```

Shows Chlorophyll a and Chlorophyll c on the x axis with site color coded.

```{r}
pig_plot+
  geom_boxplot(color="black") +
  scale_fill_manual(values=c("skyblue3","coral3"), 
                    name= NULL, labels=c("Normoxic","Hypoxic"))
```

#### Chlorophyll a (Chla) Plots only

Assign data to ggplot set a jitter object to offset points.

```{r}
jitter2 <- position_jitter(width=0.05)  
chla_plot <- ggplot(means.all,
                     aes(x=site, y=chla_mean))
```

```{r}
cp <- chla_plot+
  geom_boxplot(color="black",
               aes(fill=site),
               width=0.4) +
  scale_fill_manual(values=c("skyblue3","coral3"), name= NULL, labels=c("Normoxic","Hypoxic")) +
  geom_point(size=2, position=jitter) +
  labs(x="Site",
       y=("Chl a Concentration (ug cm^-2)")) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(colour="black"), 
        axis.text.y = element_text(colour="black"),
        axis.title.y = element_text(margin = margin(t=0, r=10,b=0,l=0)),
        panel.border = element_rect(colour = "black", fill=NA, size=.5)) +
  theme(text = element_text(size=12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay","Roldan" = "Cayo Roldan")) +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,6))
cp
```

```{r, echo=FALSE, fig.height=4, layout='l-body-outset', eval=FALSE}
bar_plots <-  yp + zp + cp
bar_plots + plot_annotation(tag_levels = 'A',
  title = 'Heatmaps of taxa distribution',
  #subtitle = 'Top 30 taxa of non-rarefied data',
  caption = 'A) DETAILS,
  B) DETAILS,
  C) DETAILS'
)
```

```{r, echo=FALSE, eval=FALSE}
plot_grid(yp, cp,zp, labels="AUTO", ncol=3, align="v")
```

### Fine Tuning

#### Maximum Quantum Yield by site

```{r}
ypp <- yield_plot +
  geom_boxplot(color = "black", aes(fill = site), width = 0.4) +
  scale_fill_manual(values = c("skyblue3", "coral3"), name = NULL, 
                    labels = c("Normoxic", "Hypoxic")) +
  geom_point(size = 1, position = jitter) +
  labs(x = "Site", y = ("Maximum Quantum Yield")) +
  theme_classic() +
  theme(legend.position = "top", legend.direction = "vertical", #
        legend.title= element_text(color = "transparent"),
        axis.text.x = element_text(size = 7, colour="black"), 
        axis.text.y = element_text(size = 8, colour = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 9, 
                                    margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0)),
        panel.border = element_rect(colour = "black", 
                                    fill = NA, size = 0.5)) +
  theme(text = element_text(size = 12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay", "Roldan" = "Cayo Roldan")) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,0.7))
```

#### Zooxanthellae Counts by site

```{r}
yzield_zoox <- ggplot(means.all, aes(x = site, y = zoox_mean))

zpp <- yield_zoox+
  geom_boxplot(color="black",
               aes(fill=site),
               width=0.4) +
  scale_fill_manual(values=c("skyblue3","coral3"), name= NULL, 
                    labels=c("Normoxic","Hypoxic")) +
  geom_point(size=1, position=jitter) +
  labs(x = "Site", y=("Endosymbiont Density (cells cm^-2)")) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 7, colour="black"), 
        axis.title.x = element_text(size = 12),
        axis.text.y = element_text(size = 8, colour="black"),
        axis.title.y = element_text(size = 9, margin = margin(t=0, r=10,b=0,l=0)),
        panel.border = element_rect(colour = "black", fill=NA, size=.5)) +
  theme(text = element_text(size=12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay","Roldan" = "Cayo Roldan")) +
  scale_y_continuous(labels = scientific,
                     breaks=seq(0,1.5E6, 4.5E5),
                     expand = c(0,0), 
                     limits = c(0,1.5E6))
```

#### Chlorophyll a (Chla)

```{r}
chla_plot <- ggplot(means.all, aes(x = site, y = chla_mean))

cpp <- chla_plot +
  geom_boxplot(color="black", aes(fill=site), width=0.4) +
  scale_fill_manual(values = c("skyblue3","coral3"), 
                    name = NULL, labels = c("Normoxic", "Hypoxic")) +
  geom_point(size=1, position=jitter) + 
  labs(x = "Site", y = ("Chl a Concentration (ug cm^-2)")) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 7, colour="black"), 
        axis.text.y = element_text(size = 8, colour="black"),
        axis.title.y = element_text(size = 9, 
                                    margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0)),
        axis.title.x = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=.5)) +
  theme(text = element_text(size=12)) +
  scale_x_discrete(labels=c("Coral" = "Coral Cay","Roldan" = "Cayo Roldan")) +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,6))
```

## Panel Figure

```{r, echo=FALSE, fig.height=5, layout='l-body-outset'}
bar_plots_ft <-  ypp + zpp + cpp
bar_plots_ft + plot_annotation(tag_levels = 'A',
  title = 'Coral response variables ',
  #subtitle = 'Top 30 taxa of non-rarefied data',
  caption = 'A) Pulse Amplitude Modulated (PAM),
  B) Zooxanthellae,
  C) Chlorophyll a (Chla)'
)
```


```{r, eval=FALSE, echo=FALSE}
panel <- plot_grid(ypp, zpp,cpp, labels="AUTO", ncol=1, align="v")
panel
```


## Source Code {.appendix}

The source code for this page can be accessed on GitHub by [clicking this link](https://github.com/hypocolypse/web/blob/master/field.Rmd).
